---
title: "p8105_hw1_xx2485.Rmd"
author: "Xiaoni Xu"
date: "2024-09-12"
output: html_document
---



## Problem 1
download the a package containing the penguins dataset:
```{r}
# install.packages("palmerpenguins")
data("penguins", package = "palmerpenguins")
library(tidyverse)
library(ggplot2)
```

Write a short description of the penguins dataset , please include:

the data in this dataset, including names / values of important variables
the size of the dataset (using nrow and ncol)
the mean flipper length

In 'penguins' dataset, there are both qualitative and quantitative data about each individual penguin. The qualitative data includes species, location, sex, and year, while the quantitative data includes bill length, bill depth, flipper length, and body mass.

The size of the dataset can be found using the code below, which shows that there are 344 individuals with 8 categories of data.
```{r}
nrow(penguins)
ncol(penguins)
```

The mean flipper length can be found using the code, which is 200.9152 after removing NA values.
```{r}
# clean the data first for the flipper length
penguins_filtered <- penguins %>%
  filter(!is.na(flipper_length_mm))

mean(penguins_filtered$flipper_length_mm)
```

A scatterplot of flipper_length_mm (y) vs bill_length_mm (x); color points using the species variable (adding color = ... inside of aes in your ggplot code should help).
```{r}
penguins_filtered <- penguins %>%
  filter(!is.na(flipper_length_mm) & !is.na(bill_length_mm))

# Create the scatterplot
scatterplot <- ggplot(data = penguins_filtered, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(title = "Flipper Length vs Bill Length by Species", 
       x = "Bill Length (mm)", 
       y = "Flipper Length (mm)")

# Display the plot
print(scatterplot)

# Export the plot to the project directory
ggsave("scatterplot_flipper_vs_bill.png", plot = scatterplot, width = 7, height = 5)
```

## Problem 2
Creating a data frame comprised of:
a random sample of size 10 from a standard Normal distribution
a logical vector indicating whether elements of the sample are greater than 0
a character vector of length 10
a factor vector of length 10, with 3 different factor “levels”

```{r}
df = 
  tibble(
    ran_samp = rnorm(n = 10),
    log_vec = ran_samp > 0,
    cha_vec = c("Calidris_canutus", "Calidris_pusilla", "Calidris_himantopus", "Calidris_subruficollis", "Calidris_alba", "Calidris alpina", "Calidris_maritima", "Calidris_bairdii", "Calidris_fuscicollis", "Calidris minutilla"),
    fac_vec = factor(c("rare", "common", "uncommon", "rare", "common", "uncommon", "rare", "rare", "rare", "common"), levels = c("common", "uncommon", "rare"))
  )
# the factor vector shows how common each Calidris species (sandpipers) is in Cook County, Illinois. Data source: my personal birding experience from 2019 to 2024 in Chicago
# the numeric values don't really have real meanings

print(df)
```


Taking the mean of each variable in your dataframe. 
```{r}
mean(df$ran_samp)
mean(df$log_vec)
mean(df$cha_vec)
mean(df$fac_vec)
```
We can take the mean for the random sample of size 10 from a standard Normal distribution and the logical vector indicating whether elements of the sample are greater than 0, but we can't take the mean for the character vector of length 10 or the factor vector of length 10.


In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?